Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-19 latex/main.tex
Update to clarify who uses MAT

Similarly to the Mutation Annotated Tree (MAT)~\cite{McBroome2021} format used 
by UShER, tskit incorporates mutational information directly in the data 
model. This is done using the site (defining the genome position and 
ancestral state) and mutation (defining the site, node, and derived state)
tables. In addition, tskit provides a flexible system for attaching metadata
to all aspects of the data model, which can either be in JSON format 
for flexibility (used heavily by sc2ts for debugging information) 
or in a vectorisable packed binary format for efficiency (used to attach
sample identifiers and pango assignments in the final ARGs). 
Tskit also provides a ``provenance'' table, which provides a mechanism
for recording software version and parameter information through each
step of complex processing pipelines.

>>>>

Similarly to the Mutation Annotated Tree (MAT)~\cite{McBroome2021} format used 
by UShER (RIPPLES~\cite{Turakhia2022} and matOptimize~\cite{DeMaio2023} are 
included in the UShER software distribution), 
tskit incorporates mutational information directly in the data 
model. This is done using the site (defining the genome position and 
ancestral state) and mutation (defining the site, node, and derived state)
tables. In addition, tskit provides a flexible system for attaching metadata
to all aspects of the data model, which can either be in JSON format 
for flexibility (used heavily by sc2ts for debugging information) 
or in a vectorisable packed binary format for efficiency (used to attach
sample identifiers and pango assignments in the final ARGs). 
Tskit also provides a ``provenance'' table, which provides a mechanism
for recording software version and parameter information through each
step of complex processing pipelines.
